### 2012-12-03 project/latex_for_presentation/test.tex
latex stuff

\documentclass[10pt,journal,letterpaper,compsoc, draftclsnofoot, onecolumn]{IEEEtran} %draft mode for double-spaced
\usepackage{graphicx}
\usepackage{subfig} % for horizontal figures
\usepackage{amsmath}
\ifCLASSOPTIONcompsoc
  % IEEE Computer Society needs nocompress option
  % requires cite.sty v4.0 or later (November 2003)
  % \usepackage[nocompress]{cite}
\else
  % normal IEEE
  % \usepackage{cite}
\fi
\usepackage{cite}

\begin{document}

blah blah i'm some math

\end{document}