### 2012-12-03 project/latex_for_presentation/test.tex
getting really close to the formatting that I want for the equation

\documentclass[10pt,journal,letterpaper,compsoc, draftclsnofoot, onecolumn]{IEEEtran} %draft mode for double-spaced
\usepackage{graphicx}
\usepackage{subfig} % for horizontal figures
\usepackage{amsmath}

%thanks to this post for \bfrac: http://www.latex-community.org/forum/viewtopic.php?f=46&t=14773
\newcommand*{\bfrac}[2]{\genfrac{(}{)}{0pt}{}{#1}{#2}} %fraction without line

\begin{document}

%$min( ($  sheddable loads $) *$ weights $+ ($ generator selection $) * $ weights $+ ($ number of generators used $) )$  %ew this one is ugly

%$min(\bfrac{sheddable}{loads} * weights +  \bfrac{generator}{selection} * weights +  \bfrac{number of}{generators used} $ %simpler version

$min(\bfrac{\text{sheddable}}{\text{loads}} * \text{\small weights} +  \bfrac{\text{generator}}{\text{selection}} * \text{\small weights} +  \bfrac{\text{number of}}{\text{generators used}} $

\end{document}